### 2008-10-08 doc/a-rex/arex_tech_doc.tex
update for new cache

The A-REX can cache input files. Caching is enabled if corresponding
command is present in configuration file. The A-REX does not cache
files marked as executable in job. Caching can also be explicitly
turned off by user for each file by using \textit{cache=no} option
in URL (for URL options read {}``Protocols, Uniform Resource Locators
(URL) and extensions supported in ARC''~\cite{urls}). The disc space
occupied by cache is controlled by removing unused files. For more
information look in Section~\ref{SubSection:ConfigFile}.

\begin{framed}
NOTE: The caching algorithm used by A-REX and described above
is currently being replaced by a better one which is no robust and
provides better scalability.
\end{framed}

@@

Cache directory contains plain files. Those are

\begin{itemize}
\item \textit{list} -- stores names of the files (8 digit numbers) and corresponding
URLs delimited by blank space. Each pair is delimited by some amount
of \textbackslash{}0 codes. Also creation and expiration times are
stored if available
\item \textit{old} -- stores URLs which have been removed from cache. Records
are delimited by some amount of \textbackslash{}0 codes and are meant
to be removed by some external routine.
\item \textit{new} -- stores URLs which have been added to cache. Records
are delimited by some amount of \textbackslash{}0 codes and are removed
when corresponding files are removed from cache. They can also be
handled by some external routines. Every time record is added to \emph{old}
it is removed from \emph{new}.
\item \textit{statistics} -- consists of strings containing \emph{name=value}
pairs. Following names are defined:

\begin{itemize}
\item \textit{hardsize} -- size of file system for storing cached data
\item \textit{hardfree} -- amount of disc space available on that file system
\item \textit{softsize} -- if cache exceeds this size files are started being
removed 
\item \textit{softfree} -- space left till softsize (can be negative)
\item \textit{claimed} -- space used by files claimed by running jobs
\item \textit{unclaimed} -- space used by files not being currently used
by any job
\end{itemize}
\item \textit{\#\#\#\#\#\#\#\#.info} -- stores state of file (\textit{\#\#\#\#\#\#\#\#}
stands for 8 digits). State is represented by one character:\\
\textit{c} -- just created, content is empty.\\
\textit{f} -- failed to download (treated same as 'c').\\
\textit{r} -- ready to be used, content is valid.\\
\textit{d} -- being downloaded. 'd' is followed by identifier of application/job
downloading that file. During content's download this file has write
lock set.
\item \textit{\#\#\#\#\#\#\#\#.claim} -- stores list of identifiers of applications/jobs
using this file. Identifiers are stored one per line.
\item \textit{\#\#\#\#\#\#\#\#} -- files storing content of corresponding
URL. These can be stored in separate directory.
\end{itemize}

Files \textit{list, old, new} and \textit{\#\#\#\#\#\#\#\#.info}
has to be stored on file system which has support for files' locking.

@@

If a job requests an input file which is subject for caching, it is
stored in the cache directory instead of the SD. The file is made
available to the job by either soft-linking it in the SD or copying
it to the SD. The latter option is more secure and hence advised.

Before downloading a file the A-REX tries to determine it's size and
then tries to preallocate space in the cache directory by writing
an empty file of the same size. If this fails (possibly because the
file system has no more space), it tries to the remove oldest cached
files which are not being used by any job. This means that \textbf{the
real hard limit of cache size is the space available in the file-system}.
In case cache gets full and it is impossible to free enough space,
the download fails and then is retried without using cache.

Before giving access to an already cached file, the A-REX contacts
the original file source to check if the user has sufficient access
rights to the file. Not all protocols make this check possible. Also,
file creation or validity times are checked to make sure the cached
file is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

The A-REX checks the cache periodically. If the space used by cache
exceeds the high water-mark given in the configuration file (\emph{softsize}),
it tries to remove the oldest unused files until cache size drops
below the low water-mark. This sets the soft limit of cache size.

There are two kinds of caches supported. Files in \textit{private}
cache are owned by the Unix user to which a Grid user is mapped. Those
files are readable only by that particular Unix user. Another kind
of cache is \textit{shared}. Files in the shared cache are owned by
the Unix user who started A-REX and are readable by everyone.

@@

\item \textbf{\textit{cacheregistration}}\textit{=yes|no} -- enables or disables
registration of cache data into Indexing Services like RC or RLS.
The default is \emph{no}. Only files downloaded through \emph{meta-url}
are registered. Registration is done to same service used for obtaining
information about file. For this operation credentials of the A-REX
(host key and certificate) are used. If required new files storage
location is registered at Indexing Service with quasi-url \emph{cache://hostname/}
and name \emph{hostname:cache}.

@@

\begin{itemize}
\item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
address \textbf{from} which the notification mails are sent.
\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
time in seconds for the SD to be available after job finished (\emph{ttl})
and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
are 7days for \emph{ttl} and 30 days for \emph{ttr}.
\item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
- specifies names for the LRMS and queue. Queue name can also be specified
in the JD (currently it is not allowed to override LRMS by usingon
the JD).
\item \textbf{\textit{session}}\textit{=path} -- specifies path to the directory
in which the SD is created. If the path is {*} the default one is
used -- \textit{\$HOME/.jobs}. In new configuration file this command
is called \textbf{\textit{sessiondir}}.
\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
the directory to store cached data. Empty \emph{path} disables caching.
Default is not to cache data. Optional \textit{link\_path} specifies
the path at which cache is accessible at computing nodes. If \textit{link\_path}
is set to '.' files are not soft-linked, but copied to session directory.
\item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} -- same
as \textit{cache} command, but cache belongs (owned) to user. For
shared caches use 'cache'.
\item \textbf{\textit{cachedata}}\textit{=path} -- allows to specify separate
place to store cache files containing data itself. This can be useful
in case of big data storage available only on NFS server which does
not support file locking. If command or \textit{path} is missing -
default is to store data at place specified in \textit{cache} or \textit{privatecache}
command, together with control files. 
\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} -- specifies
high and low water-mark for space used by cache. Values are specified
in bytes. Both \textit{high\_mark} and \textit{low\_mark} can be negative
values. In that case corresponding positive value means space left
on file system. If \textit{low\_mark} is omitted it becomes equal
to \textit{high\_mark}. By default this feature is turned off. To
turn it off explicitly \textit{cachesize} without parameters should
be specified. If turned off cache will grow up till it fills whole
file system.
\item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number
of times job will be allowed to rerun after it failed at any stage.
Default value is \emph{5}. This only specifies a upper limit. Actual
number is provided in job description and defaults to 0.
\end{itemize}

>>>>

The A-REX can cache input files. Caching is enabled if one or more cache
directories are specified in the configuration file. The A-REX does not
cache files marked as executable in a job. Caching can also be explicitly
turned off by the user for each file by using the\textit{ cache=no}
option in the URL (for URL options read {}``Protocols, Uniform Resource
Locators (URL) and extensions supported in ARC'' \cite{urls}). The
disk space occupied by the cache is controlled by removing old files.
For more information look in section \ref{SubSection:ConfigFile}.

@@

Cached files are stored in sub-direcories under the \emph{data} directory
in each main cache directory. Filenames are constructed from an SHA-1
hash of the URL of the file, and split into subdirectories based on
the two initial characters of the hash. This enables the cached files
to be evenly split over a number of subdirectories. If
more than one cache directory is used the initial letter of the hash
also determines which cache is to be used. The algorithm is simply
to use the cache directory at index \emph{i} in the list of directories,
where \emph{i} is found from the initial letter of the hash mod the
number of caches. This limits the maximum number of cache directories
to 16, as SHA-1 hashes use the character set {[}\emph{0-9,a-f}]. In
the extremely unlikely event of a collision between two URLs having
the same SHA-1 hash, caching will not be used for the second file.

There is no indexing system for the cache, and any cache filename
can be easily determined from a URL by using the same hashing algorithm
as the A-REX, e.g. the standard command line tool \emph{sha1sum. }The
cache directory used by a file can be determined from the initial
letter of the hash and the order of cache directories in the configuration
file. Some associated metadata (the corresponding URL and an expiry
time, if available) are stored in a file with the same name as the
cache file, with a \emph{.meta} suffix.

For example, with a cache directory \emph{/cache}, the file

\begin{center}
\emph{lfc://atlaslfc.nordugrid.org//grid/atlas/file1 }\\
is mapped to\\
\emph{/cache/data/78/f607405ab1df6b647fac7aa97dfb6089c19fb3}
\par\end{center}

and the file \emph{/cache/data/78/f607405ab1df6b647fac7aa97dfb6089c19fb3.meta
}contains the original URL and an expiry time if one is available.

At the start of a file download, the cache file is locked, so that
it cannot be deleted and so that another download process cannot write
the same file simultaneously. This done by creating a file with the
same name as the cache filename but with a \emph{.lock} suffix. This
file contains the process ID of the process and the hostname of the
host holding the lock. If this file is present, another process cannot
do anything with the cache file and must wait until the cache file
is unlocked (i.e. the \emph{.lock} file no longer exists). The lock
has a timeout of one day, so that stale locks left behind by a download
process exiting abnormally will eventually be cleaned up. Also, if
the process corresponding to the process ID stored inside the lock
is no longer running on the host specified in the lock, it is safe
to assume that the lock file can be deleted.

@@

If a job requests an input file which can be cached or is allowed to
be cached, it is stored in the selected cache directory, and depending
on the configuration, either the file is copied to the SD or a hard
link is created in a per-job directory and a soft link is created in
the SD to there. The per-job directories are in the \emph{joblinks}
subdirectory of the main cache directory. The former option is advised
if the cache is on a file system which will suffer poor performance
from a large number of jobs reading files on it, or the file system
containing the cache is not accessible from worker nodes. The latter
option is the default option. The per-job directory is only readable
by the local user running the job, and the cache directory is readable
only by the A-REX user (usually root). This means that the local user
cannot access any other users' cache files. It also means that cache
files can be removed without needing to know whether they are in use
by a currently running job.

If the file system containing the cache is full and it is impossible
to free any space, the download fails and is retried without using
cacheing.

Before giving access to a file already in the cache, the A-REX contacts
the initial file source to check if the user has read permission on
the file. Also file creation or validity times from the original source
are checked to make sure the cached file is fresh enough. If it is
impossible to obtain creation and invalidation times for the file,
it is invalidated 24 hours after download.

The A-REX checks the cache periodically. If the used space on the file
system containing the cache exceeds the high water-mark given in the
configuration file it tries to remove the least-recently accessed
files to reduce size to the low water-mark.

\subsection{Administration tools}

The following tools (installed in \emph{\$ARC\_LOCATION/libexec})
exist to help with administration of the cache:

\begin{itemize}
\item \emph{cache-clean} - This tool is used periodically (every 2 minutes)
by the A-REX to keep the size of each cache within the configurated limits.
It removes files from the cache if the total size of the cache is
greater than the configured limit. It will attempt to remove files
which are not locked in order of access time, starting with the earliest,
until the size is lower than the configured lower limit. If the lower
limit cannot be reached (because too many files are locked, or other
files outside the cache are taking up space on the file system), the
tool will exit before the lower limit is reached.\\
\emph{cache-clean -h} gives a list of options. The most useful option
for administrators is \emph{-s}, which does not delete anything, but
gives summary information on the files in the cache, including information
on the ages of the files in the cache.\\
It is not recommended to run \emph{cache-clean }manually to clean
up the cache, unless it is desired to temporarily clean up the cache
with different size limits to those specified in the configuration.
\item \emph{cache-list} - This tool is used to list all files present in
each cache. It simply reads through all the \emph{.meta }files and
prints to stdout a list of all URLs stored in each cache and their
corresponding cache filename, one per line.
\end{itemize}

@@

@@

\begin{itemize}
\item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
address \textbf{from} which the notification mails are sent.
\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
time in seconds for the SD to be available after job finished (\emph{ttl})
and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
are 7days for \emph{ttl} and 30 days for \emph{ttr}.
\item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
- specifies names for the LRMS and queue. Queue name can also be specified
in the JD (currently it is not allowed to override LRMS by usingon
the JD).
\item \textbf{\textit{session}}\textit{=path} -- specifies path to the directory
in which the SD is created. If the path is {*} the default one is
used -- \textit{\$HOME/.jobs}. In new configuration file this command
is called \textbf{\textit{sessiondir}}.
\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
a directory to store cached data (see \ref{sec:cache}).\emph{ }Multiple
cache directories may be specified by specifying multiple \emph{cachedir}
commands. Cached data will be distributed evenly over the caches.
Specifying no \emph{cachedir }command or commands with an empty path
disables caching. Optional \textit{link\_path} specifies the path
at which \emph{path} is accessible on computing nodes, if it is different
from the path on the A-REX host. If \textit{link\_path} is set to '.'
files are not soft-linked, nor are per-job links created, but files
are copied to the session directory. In old configuration file this
command is called \textbf{\textit{cache}}.
\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies
high and low watermarks for space used by each cache, as a percentage
of the space on the file system on which the cache directory is located.
When the max is exceeded, files will be deleted to bring the used
space down to the min level. It is a good idea to have each cache
on its own separate file system. To turn off this feature, \char`\"{}cachesize\char`\"{}
without parameters can be specified. These cache settings apply to
all caches specified by \emph{cachedir} commands.
\item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number
of times job will be allowed to rerun after it failed at any stage.
Default value is \emph{5}. This only specifies a upper limit. Actual
number is provided in job description and defaults to 0.
\end{itemize}
